---
title: "Assignment 7th"
author:
  name: Maksymilian Gmur
  date: 04.25.2025
  affiliations:
    id: JU
    name: Jagiellonian University
format: 
  html:
      df-print: paged
      tidy: false
      embed-resources: true
      theme: lumen
      toc: true
      toc-location: left
      toc-depth: 6
      toc-expand: 6
      code-tools:
        source: true
        toggle: true
      code-fold: false
      code-summary: "Show the code"
      code-overflow: wrap
      page-layout: full
editor: visual
editor_options: 
  markdown: 
    wrap: 72
execute:
  cache: true
---

# Task #1

Firstly I will start with loading all essential libraries.

```{r}
library(tidyr)
library(dplyr)
library(readr)
library(datasets)
```

The code block below was created with help of ChatGPT, to work around malfunctioning of ***parsed_table()*** function

```{r}
install.packages("DT")
library(DT)
```

Progressing further I will solve task number one, with creation of *i**ris_long*** table. To do that, the pivot_longer function was used. I chose which columns should be converted with ***cols=match*** and then specify columns where the names and values of those columns will go. That is how I will obtain dataset with 3 columns: ***Species \| Trait \| Size***

```{r}
data("iris")
View(iris)

iris_long = pivot_longer(iris, cols=matches("Sepal|Petal"), names_to="Trait", values_to="Size" )
```

# Task #2

## #2.1

To compute mean values of specific traits in iris dataset I will use pipeline which will:

1.  Group rows by paired values of Species and Traits columns, creating distinctive groups.

2.  Those distinctive groups are then used in the input of summarize function, which summarize all mean values for each group.

```{r}
iris_mean <- iris_long %>%
  group_by(Species, Trait) %>%
  summarize(MeanValue = mean(Size))
DT::datatable(iris_mean)
```

Output table contains all mean values for each trait, according to different species.

## #2.2

This task was pretty straightforward. It is very similar to previous task, however here I had to use ***filter()*** and ***max()*** functions additionally.

```{r}
iris_long %>%
  filter(Species %in% c("virginica","versicolor")) %>%
  group_by(Species, Trait) %>%
  summarize(MaxValue = max(Size))
```

## #2.3

This one was fun to solve. Basically, I thought about filtering ***Size column*** values through comparison to ***MeanValue column***. To do that, I used ***left_join()*** function, so every row would have all the appropriate values. Here I made iris_test table to check if my code worked correctly. After that was proved, I just added ***group_by()*** function to, well, group specific rows and then counted them using ***n()*** function. To be honest, I don't know how exactly this function works, so I checked on the internet that people use it as presented below. I tested it and here I had, probably, the correct answer. 

```{r}
iris_test <- iris_long %>%
  left_join(y=iris_mean, by = c("Species", "Trait")) %>%
  filter(Size > MeanValue, .by = Size)
DT::datatable(iris_test)
```

```{r}
iris_combined <- iris_long %>%
  left_join(y=iris_mean, by = c("Species", "Trait")) %>%
  filter(Size > MeanValue, .by = Size) %>%
  group_by(Species, Trait) %>%
  summarize(Count = n())
DT::datatable(iris_combined)
```
